---
title: "EDA"
author: "Enzo Moraes Mescall"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(lme4)
library(lmerTest)
```

```{r message=FALSE, warning=FALSE}
# Importing data
path = "./data/SC - Endorsment Data Analysis - "
precinct_df = read_csv(paste0(path, "precinct.csv"))
candidates_df = read_csv(paste0(path, "candidates.csv"))

aj_v_leo = read_csv(paste0(path, "AJ vs. Leo.csv"))
baker_v_caballero = read_csv(paste0(path, "Baker vs. Caballero vs. Rist vs. et. al.csv"))
johnson_v_reece = read_csv(paste0(path, "Johnson vs. Reece vs. Caballero vs. et. al.csv"))
murdock_v_freelon = read_csv(paste0(path, "Murdock vs. Freelon.csv"))
satana_v_echols = read_csv(paste0(path, "Satana vs. Echols (County Wide).csv"))
schewel_v_ali = read_csv(paste0(path, "Schewel vs. Ali.csv"))
```

## Methodology

Build a table which looks like this:

Result | Precinct | [Candidate 1 info ] | [Candidate 2 info] | [Population info]

Where results is the gap between the to candidates. In a multi-candidate election we can artificially turn it into a series of head-to-head races. We then want to regress on result as the dependent variable and identify the independent variables that are most predictive of the result.

```{r}
# Buliding the master table

# Demographic percentages of possible voters in the precinct
precinct_pct = data.frame(lapply(precinct_df %>% select(-precinct_id), function(x) x/precinct_df$Total)) %>%
  mutate(row = row_number()) %>%
  right_join(precinct_df %>% select(precinct_id) %>% mutate(row = row_number()), by = "row") %>%
  select(-row, -Total) %>%
  select(precinct_id, everything()) %>%
  rename("Precinct" = "precinct_id") 
```

```{r}
pct_to_dec = function(string) {
  string = gsub("%", "", string)
  string = as.numeric(string)
  string = string/100
  return(string)
}


# Creating a function to clean up race dataframes
head_to_head = function(candidate_df, candidate_names){
  # Initialize hth_df with Precinct, c1_name, c2_name, result, c1_pct, c2_pct
  hth_df = data.frame(Precinct = character(),
                      c1_name = character(),
                      c2_name = character(),
                      result = numeric(),
                      c1_pct = numeric(),
                      c2_pct = numeric())
  
  # Create a matrix of all combination of candidate_names
  candidate_combinations = expand.grid(candidate_names, candidate_names) %>%
    filter(Var1 != Var2) %>%
    mutate(Var1 = as.character(Var1), Var2 = as.character(Var2))
  
  # Loop through the combinations of two names and create a new dataframe
  for(i in 1:nrow(candidate_combinations)){
    candidate_1 = candidate_combinations$Var1[i]
    candidate_2 = candidate_combinations$Var2[i]
    
    hth = candidate_df %>%
      mutate(c1_pct = pct_to_dec(get(paste0("%", candidate_1))),
             c2_pct = pct_to_dec(get(paste0("%", candidate_2))),
             result = c1_pct - c2_pct,
             c1_name = candidate_1,
             c2_name = candidate_2) %>%
      select(Precinct, c1_name, c2_name, result, c1_pct, c2_pct)
    
    # Append hth to the bottom of hth_df
    hth_df = bind_rows(hth_df, hth)
  }
  return(hth_df)
}
```

```{r message=FALSE, warning=FALSE}
aj_v_leo_hth = head_to_head(aj_v_leo, c("AJ Williams", "Leonardo Williams"))
baker_v_caballero_hth = head_to_head(baker_v_caballero, c("Nate Baker", "Javiera Caballero1", "Carl Rist", "Khalilah Karim", "Monique Holsey-Hyman", "Shelia Ann Huggins"))
johnson_v_reece_hth = head_to_head(johnson_v_reece, c("Jillian Johnson", "Charlie Reece", "Javiera Caballero0", "Daniel Meier", "Jacqueline Wagstaff"))
murdock_v_freelon_hth = head_to_head(murdock_v_freelon, c("Natalie Murdock", "Pierce Freelon", "Gray Ellis"))
satana_v_echols_hth = head_to_head(satana_v_echols, c("Satana Deberry", "Roger Echols", "Daniel Meier"))
schewel_v_ali_hth = head_to_head(schewel_v_ali, c("Steve Schewel", "Farad Ali"))
```

```{r}
# binding all
hth_df = bind_rows(aj_v_leo_hth, baker_v_caballero_hth, johnson_v_reece_hth, murdock_v_freelon_hth, satana_v_echols_hth, schewel_v_ali_hth)
```

```{r}
# Joining with precinct_pct and candidates_df
full = hth_df %>%
  left_join(candidates_df %>% select(-id), by = c("c1_name" = "name")) %>%
  left_join(candidates_df %>% select(-id), by = c("c2_name" = "name"), suffix = c(".c1", ".c2")) %>%
  left_join(precinct_pct, by = "Precinct")  %>%
  filter(Precinct != "ABSENTEE" | Precinct != "ABS")

only_sophia = hth_df %>%
  left_join(candidates_df %>% select(-id), by = c("c1_name" = "name")) %>%
  left_join(candidates_df %>% select(-id), by = c("c2_name" = "name"), suffix = c(".c1", ".c2")) %>%
  inner_join(precinct_pct, by = "Precinct") 
  
# write.csv(full, "./data/full_elections.csv")
# write.csv(only_sophia, "./data/full_elections.csv")
```


```{r}
# Making key variables to regress on 


# create candidate precinct alignment

```

## Model


TODO WIP
Let $y_1,\dots,y_n$ be independent random variables such that $y_i \sim Beta(\mu, \phi)$ where $E(y) = \mu$ and unknown precision $\phi$. The logit beta regression model will rely on the following parameters:

* $y_i$ is the observed percentage of a head-to-head race between candidates $c_1$ and $c_2$ in precinct $i$.
* Fixed effect coefficients:
  + $\beta_0$ is the intercept
  + $\beta_1$ for whichever endorsement we are interested in
  + $\beta_2$ for `incubent`
  + $\beta_3$ for interaction between `race` and `race.c1``
  + $\beta_4$ for interaction between `race` and `race.c2``
  + $\beta_5$ 
  + $\beta_6$ 
* Random effects:
  $b_1(\text{Precinct}_i)$ is the random intercept for the i-th precinct
   $\(b_2(\text{Precinct}_i) \times \text{Age}_{j[i]}\)$ is the random intercept for the j-th age group nested within the i-th precinct.
Finally, $\epsilon_{ij}$ represents the residual error for the j-th age group in the i-th precinct.


```{r}
# family = list(family="beta",link="logit")

model = lmer(result ~ -1 + PA.c1 + Indy.c1 + Committee.c1 + PA.c2 + Indy.c2 + Committee.c2 + age.c1 + sex.c1 + (-1 + Indy.c1 + PA.c2 + Committee.c2 + race.c1 + age.c1 + sex.c1|Precinct), data = only_sophia)
summary(model)
```
```{r}
precinct_effects = ranef(model)
fixed_effects = summary(model)$coefficients
```

```{r}
sophia_pct = precinct_effects$Precinct %>% 
  mutate(Indy.c1 = round((Indy.c1 + fixed_effects[2])/2, 3),
         PA.c2 = round((PA.c2 + fixed_effects[4])/2, 3),
         Committee.c2 = round((Committee.c2 +fixed_effects[6])/2, 3),
         age.c1old = round((age.c1old + fixed_effects[8])/2, 3),
         sex.c1male = round((sex.c1male + fixed_effects[10])/2, 3)) %>%
  select(Indy.c1, PA.c2, Committee.c2, age.c1old, sex.c1male) %>% 
  mutate(total = (Indy.c1 + PA.c2 + Committee.c2 + age.c1old + sex.c1male)/5)

write.csv(sophia_pct, "./data/precinct_all_factors_pct.csv")
```

```{r}
precinct_csv =  sophia_pct %>%
  mutate(effects = Indy.c1 + PA.c2 + Committee.c2 + age.c1old + sex.c1male) %>% select(effects) %>%
  mutate(effects = (effects - min(effects)) / (max(effects) - min(effects))) %>%
  rename(precinct_importance = effects)

write.csv(precinct_csv %>% mutate(precinct_importance = round(precinct_importance, 3)), "./data/precinct_all_factors_index.csv")
```

```{r}
endorsements_pct = precinct_effects$Precinct %>%
  mutate(Indy = (Indy.c1 + fixed_effects[2])/2, PA = (PA.c2 + fixed_effects[4])/2, Committee = (Committee.c2 +fixed_effects[6])/2) %>%
  select(Indy, PA, Committee) %>%
  mutate(total = (Indy + PA + Committee)/3)

write.csv(endorsements_pct %>% mutate(Indy = round(Indy, 3), PA = round(PA, 3), Committee = round(Committee, 3)), "./data/precinct_endorsements_pct.csv")
```

```{r}
endorsements = endorsements_pct %>%
  select(Indy, PA, Committee) %>%
  mutate(Indy = (Indy - min(Indy)) / (max(Indy) - min(Indy)),
         PA = (PA - min(PA)) / (max(PA) - min(PA)),
         Committee = (Committee - min(Committee)) / (max(Committee) - min(Committee)))

write.csv(endorsements %>% mutate(Indy = round(Indy, 3), PA = round(PA, 3), Committee = round(Committee, 3)), "./data/precinct_endorsements_index.csv")
```

