---
title: "EDA"
author: "Enzo Moraes Mescall"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
# Importing data
path = "./data/SC - Endorsment Data Analysis - "
precinct_df = read_csv(paste0(path, "precinct.csv"))
candidates_df = read_csv(paste0(path, "candidates.csv"))

aj_v_leo = read_csv(paste0(path, "AJ vs. Leo.csv"))
baker_v_caballero = read_csv(paste0(path, "Baker vs. Caballero vs. Rist vs. et. al.csv"))
johnson_v_reece = read_csv(paste0(path, "Johnson vs. Reece vs. Caballero vs. et. al.csv"))
murdock_v_freelon = read_csv(paste0(path, "Murdock vs. Freelon.csv"))
satana_v_echols = read_csv(paste0(path, "Satana vs. Echols (County Wide).csv"))
schewel_v_ali = read_csv(paste0(path, "Schewel vs. Ali.csv"))
```

## Methodology

Build a table which looks like this:

Result | Precinct | [Candidate 1 info ] | [Candidate 2 info] | [Population info]

Where results is the gap between the to candidates. In a multi-candidate election we can artificially turn it into a series of head-to-head races. We then want to regress on result as the dependent variable and identify the independent variables that are most predictive of the result.

```{r}
# Buliding the master table
```

